### 2014-03-11 paper.tex
Updated intro a little

\documentclass[twocolumn]{paper}
\usepackage{authblk}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[margin=1.0in]{geometry}
\lstset{language=Haskell}

\title{Maybe Slipstreams}
\author[*]{Morgan McDermott}
\author[*]{John Carlyle}
\affil[*]{University of flightless dragons}

\begin{document}
\maketitle

\section{Abstract}
We give a definition of \textit{Functional Reactive Programming} (FRP) that allows for dynamic signal networks and safe interaction with the real world. We allow signal networks to be dynamic by modeling them as higher-order streams (as in Patai \cite{HighOrderStreams}), and use the concept of wormholes (from Winograd \cite{WinogradCort2012HS}) to make interaction with real world resources safe. All signal values are modeled as $Maybe$ values to allow for defined behavior in the face of disrupted signals, and to reference signal values before they have any reasonable value. We also specify a DSL that uses this FRP model, allowing a programmer to easily design signal networks to accomplish tasks by specifying layers and how the layers interact. Our implementation builds on previous work by being more practical. The DSL arises naturally from our definition of FRP, is simple to use and requires only limited knowledge of the underlying FRP concepts.

\section{Introduction}
\subsection{What is FRP}
Functional Reactive Programming, introduced by Fran \cite{ElliottHudak97:Fran}, is a programming paradigm based on the key concepts of time varying values.   These values can also be considered events depending on the granularity of the stream. Reactivity is defined as a reaction to an event; when a node recieves a new value it reacts by changing its value. FRP is particularly well suited for transforming data and data structures since it is, by nature, a collection of pure functions.

\subsection{Why use FRP}
FRP is a useful approach to user interface design because it is declaritive. Most interface design forces you to define exactly how to generate each value. With the time varying values supplied by an FRP network the programmer can choose \textit{what} an interface element is rather than how it is generated. By comosing these values together a higher-level description of what each

\subsection{Problems with FRP}
We solve the problem of interacting with time-varying values can be solved using stream-based FRP, where each computation sends a stream instead of a signal value. A stream is essentially a complete history of the values of a particular computation. Any other computation that needs the current value, or a value from the past, can look at the stream and find the appropriate values. This leads to additional problems like memory usage growing out of control as new values are continually added to the stream and not removed. In some systems garbage collection finds and removes values from streams that are no longer dependencies for other computations. In Elerea \cite{HighOrderStreams}, this is solved by only allowing access to the most recent values in streams.

\subsection{Organization}
This paper primarily introduces a way for FRP to interact safely in an environment filled with hazerdous resources which are difficult to model as a function of time. (What the different sections are for) (Section 1) Section 2 Section 3 Section 4 Section 5 Section 6 Section 7.

\section{A Motivating Example}
Adapting FRP for javascript development would be super fantastic since javascript GUI code deals with varying values all the time. In particular the asynchronous communication between a server and client through the use of AJAX can be thought of as a time-varying signal between two nodes.

There are several problems that must be solved in stream-based FRP to allow this transaction to take place. First, how do we interact with the environment itself? Since we need to manipulate the DOM and perform asynchronous requests, we have to change state.

Secondly, how do we model the changing requirements for interfaces? If the user presses a button to submit a query to the server, the server could respond in a way indicating that they need to update the interface on the page. Perhaps we need to create more buttons that could perform different server queries. In order to solve this, we must define a flexible, dynamic FRP network that can modify itself. Elerea solves this by providing Higher-Order Streams.

We need to ensure that the time-based values from the server are treated in a well-defined way. Namely, what if we have never contacted the server? What should the value of its stream be, and how should that differ from the stream after it has responded and is in an idle state again? We solve this by encapsulating every stream value in a $Maybe$, as we will explain later.

Lastly, what happens when two computations wish to change the state of the interface at the same time? How can we allow safe interaction between external state and multiple FRP calculations that wish to modify the state? We solve this using Wormholes as defined by Winograd-Cort \cite{WinogradCort2012HS}.

\section{Terminology}
\begin{description}
\item[element] Any value or computation that relies upon previous computations or values. An element is analogous to a node in a graph. The element is connected to various other elements by means of some function. All elements connected to element $e$ can be split into two sets $s_1$ (all elements supplying input to $e$) and $s_2$ (all elements recieving their input from $e$). $s_1$ is called the predecessor set of $e$ or $s_1 = pre(e)$, and $s_2$ is the successor set of $e$ or $s_2 = suc(e)$

\item[network] A set of elements that form a connected component. Multiple networks can be used to describe different unrelated components of a particular interface or problem space.

\item[signal] It is useful to think of the values being sent from one element to another as a signal being broadcast continuously. The element receiving this signal adjusts its own outgoing signal according to a predefined set of rules or a function.

\item[stream] An element may have need of signals from more than one timestep in the past. A stream is essentially a history of all previous values of a signal, with the most current signal value at the head of the list. Streams travel between elements, thus allowing an element to select what timestep it selects a value from. The downside of streams is that they allow for spacetime leaks, which are described below. A stream of type $Stream~a$ will be denoted as $\langle v_0~v_1~v_2~...\rangle$ where $v_i$ is of type $
a$.

\item[behavior] Using the previous definition of an element and a signal, an element can be thought of as a function that maps from one signal to another. More concretely \\
  $type~element~a~b = Signal ~a \rightarrow Signal ~b$\\
  Behaviors are typically thought of in terms of continuous functions. What FRP is trying to do is model the change in behavior of some element in a network. This takes up a lot of processor power  and is not very efficient.

\item[event] An event is a discrete behavior. Since behaviors perform so poorly, it is often necessary to break a signal into a set of discrete values. The cutoff is not clear between high and low sampling rates. With a high sample rate, something can be considered a behavior; with a low sample rate it can be considered an event. Our implementation of FRP is skewed toward events.

\item[layer] A layer is a set of elements in a network where the intersection of their mutual union of predecessor and successor sets is the empty set. More formally $\bigcup_{e_i \in L}{pre(e_i)} \cap \bigcup_{e_j \in L}{suc(e_j)} = \emptyset$ where $L$ is a layer. Breaking a network into layers is helpful because it helps identify dependencies in the network. It also helps introduce structure into the environment, which can help organize a network which may have been difficult to think about. It is not always possible to adhere to the definition of a layer, or is not practical to do so (even if it could) be done. For example, a feedback loop would be, by definition, a loop on the network, some way of passing a result back to a predecessor. This technique is useful for maintaining state in a network amongst other uses.

\item[space-time leak] In a functional reactive service, any particular element can depend on values that are far in the past. Clearly, the longer the program is executing, the longer the history of events, or the longer the stream is. A longer stream takes up more memory and can cause slowdown if the stream is not trimmed after a certain point. Typically a garbage collection service of some kind is used to clean up old events upon which are no longer being depended.. This is analogous to a memory leak in imperative programming. The value is said to be a space-time leak if it is unnecessarily being used for computations when it will have no effect on any current elements.
\end{description}

\section{Our contributions}
Our contributions are basically jsut reworking the bind in a stream monad to map between layer sets. Also each value is wrapped in a maybe, revolutionary. This allows safe access to resources that have not yet been created in a dynamic network setting. It also allows for safe handling of volatile resources such as HTTP requests or accessing other error prone procedures. Convenience elements have been added such as the aggregator, which allows for dynamic seeming networks without any modification to the underlying structure. This is possible due to the addition of the maybe value to every stream.
\subsection{DSL Grammar}
\footnotesize
\begin{align*}
  Statement ::=& ~V \leftarrow Expression &\text{Variable assignment}\\
  |& ~Expression &\text{Single expression}\\
  Expression ::=& ~SignalNet &\text{A signal network}\\
  |& SimpleExpr &\text{An expression}\\
  SignalNet ::=& ~( Exprs ) \rightarrow SignalNet &\text{Compose layers}\\
   |& ~(Exprs ) &\text{Single layer} \\
  Exprs ::=& ~SimpleExpr ~,~ Exprs &\text{List of expressions}\\
  |& ~SimpleExpr &\text{Single expression}\\
  SimpleExpr ::=& ~L[``Javascript code''] &\text{Lift}\\
  |& ~ @V &\text{Get}\\
  |& ~ *V(SimpleExpr) &\text{Put}\\ 
  |& ~ V( Exprs ) &\text{Function call}\\
  V ::=& ~Word~Characters &\text{Variable}
\end{align*}
\normalsize
As you can see our grammar is exteremly simple and contributes basically 1 monad with bind. Thats it. PhD Please.

\subsection{Maybe Streams}
Merging streams and wormholes and making an implementation of the whole thing practical. Making it based in theory so its not a giant mess like bacon.js. Monad streams woo!

\subsection{Higher-order streams}
Streams composed with streams oh my. Gotta look out for those undefined values. The value of composed streams is the diagonal of the matrix formed by the stream values. Boop:

\begin{align*}
  \langle &\langle \mathbf{v_{00}} ~ v_{01} ~ v_{02} ~ v_{03} ~ ... ~ \rangle \\
  &\langle \perp ~ \mathbf{v_{11}} ~ v_{12} ~ v_{13} ~... ~ \rangle \\
  &\langle \perp ~ \perp ~ \mathbf{v_{22}} ~ v_{23} ~ ...  ~ \rangle \rangle
\end{align*}

Better idea to avoid wasted space and complex ge

\begin{align*}
  \langle &\langle \mathbf{v_{00}} ~ v_{01} ~ v_{02} ~ v_{03} ~ ... ~ \rangle \\
  &\langle  \mathbf{v_{11}} ~ v_{12} ~ v_{13} ~ v_{14} ~ ... ~ \rangle \\
  &\langle  \mathbf{v_{22}} ~ v_{23} ~ v_{24} ~ v_{25} ~ ...  ~ \rangle \rangle
\end{align*}

\section{Related work}

\subsection{Higher-order streams}
Work has already been done on higher-order streams \cite{HighOrderStreams}. Higher-order streams are nessisary for having a dynamic network. Dynamic networks are required

\subsection{Wormholes}
\cite{WinogradCort2012HS}
\section{Future work}
Probably some since we want to keep working on things. I bet it will have to do with extending the language definition and possibly defining streams better.

\bibliography{paper}
\bibliographystyle{plain}
\end{document}